Final paper: Towards Proving Termination of \texttt{zelda-mosaic} Algorithms under Conditions on the Input

### 2021-05-03 paper/paper.tex
paper: intro/related work

\section{Summary of Proposal}

@@

Our original goal was to examine these algorithms as implemented in the \matlab\
source and prove, if possible, their termination given well-conditioned input.
With some of the work now behind us, we have a better idea of what this project
entails and how we will be able to reach the finish line.

\section{Problem Description}

>>>>

\section{Introduction}

@@

Our original goal was to examine these algorithms as implemented in the \matlab\
source and prove, if possible, their termination given well-conditioned input.
Thus we embarked on developing a Coq model of the mosaic algorithms in order to
write proofs about them. As we will explain, we did not complete this
objective---instead, we made progress on the core of the problem: defining
matrices and their operations. To that extent, we will gloss over the finer
details of \matlab\ where possible and focus specifically on the relevant matrix
operations.

In the subsequent sections, we cover related work (\S~\ref{S:related_work}),
describe the problem we tackled (\S~\ref{S:problem_desc}), our approach to
designing a solution (\S~\ref{S:design}), and our results (\S~\ref{S:results}).

\section{Related Work}\label{S:related_work}

Because we are proving something about code that we have written ourselves,
there is not a vast body of literature specific to our algorithms from which we
can draw. There are however a number of related issues exposed by our code,
which we will discuss here. In particular, the following are of interest and
have been written about:

\begin{itemize}
    \item The correctness of \matlab's \textit{cell array} data structure and
        its representation in Coq;
    \item The correctness of \matlab's built-in numerical functions and matrix
        operations; and
    \item Proving termination of a particular program based on the inputs of the
        program.
\end{itemize}

We are not the first group to demonstrate interest in proving properties about
matrix representations. In the early days of Coq, a
paper~\cite{magaud:hal-00955444} was published discussing the implementation of
square matrices in Coq, as well as proofs involving basic operations such as
addition and multiplication. This paper references \textit{Linear
Algebra}~\cite{lin_alg}, a Coq library written by a colleague which formalizes a
well-known linear algebra text~\cite{linear_algebra} and supports matrices with
a differing number of rows and columns. These implementations are specific to
two-dimesional matrices, as far as we can tell, and are thus inadequate for the
manipulation of higher dimensions in our algorithms.

\matlab\ is a closed-source piece of software, and little has been published about
its correctness. In the official \matlab\ forum, users have demonstrated interest
in verifying the correctness of \matlab\ algorithms~\cite{duenisch_2013} and the
\matlab\ staff reassure users that the algorithms have gone through rigorous
testing, but this assurance does not carry the same weight as a formal
verification. Indeed, as we have learned, these techniques fall on separate
parts of a correctness spectrum, with testing demonstrating correctness less
reliably than formal methods. The lack of related work in the \matlab\ domain
served as further motivation to obtain results about the correctness of the subset
of \matlab\ features used in our code.

Program termination is a topic of wide interest in the computing community, as
it has both practical and of theoretical relevance given the famous halting
problem. \citeauthor{Cook_2011}~\cite{Cook_2011} argue in their article that
computer scientists are often hesitant to make claims about the termination of a
particular program, but that it is often achievable. Their work draws from the
field of compilers and makes the argument that one can prove the termination of
a program given certain inputs, provided that there is sufficient analysis of
the relation between the inputs and the possible paths that may be taken during
execution\footnote{In many contexts, this involves things like symbolic
execution (cf. Rosette~\cite{Rosette}).}. Many of the ideas in this work are
formalized by \citeauthor{BLANQUI_2011} in the \colorlib\ library for Coq,
which is used primarily for proofs involving rewriting code and program
termination~\cite{BLANQUI_2011}.

\section{Problem Description}\label{S:problem_desc}
